---
title: "1_Introduction"
output: html_document
date: "2024-07-09"
---

## 1.1. Hello, world!

Print out a given character string with the `cat` function:

```{r}
cat("My hovercraft is full of eels.\n") # \n == newline
```
### Exercise 1.1
```{r}
cat("This is the first exercise of the textbook.\n")
cat("Today I ate homemade pannacotta.")
```
## 1.3. Atomic vectors at a glance

In R, no separate types representing individual values have been defined. Instead, what seems to be a single datum, is already a vector (sequence, array) of length one.

```{r}
2.71828          # input a number; here: the same as print(2.71828)
length(2.71828)  # it is a vector with one element
```
To create a vector of any length, we can call the `c` function, which combines given arguments into a single sequence:

```{r}
c(1, 2, 3) # three values combined
length(c(1, 2, 3)) # # indeed, it is a vector of length three
```
## Exercise 1.6
The three most important atomic types are numeric, logical and character.

## Exercise 1.7
Each scalar value in R is actually a one-element vector. Atomic vectors are thus considered basic data types in R.





