---
title: "2_Numeric vectors"
output: html_document
date: "2024-07-09"
---

## 2.1. Creating numeric vectors

### 2.1.1. Numeric constants

```{r}
-3.14
123e-4 # scientific notation
NA_real_  # numeric NA (special indicator for missing value)
```
### 2.1.2. Concatenating vectors with `c`

```{r}
c(1, 2, 3)  # three vectors of length one –> one vector of length three
c(1, c(2, NA_real_, 4), 5, c(6, c(7, Inf)))
```
### 2.1.3. Repeating entries with `rep`

Replicate elements in vector.

```{r}
rep(x=1, times=5)
rep(c(1, 2, 3), 4)
# each element is repeated the corresponding number of times
rep(c(1, 2, 3), c(2, 1, 3)) # both vectors must be of same length
rep(c(1, 2, 3), c(4, 4, 4))
rep(c(1, 2, 3), each=3)
rep(c(1, 2, 3), length.out=7)
rep(c(1, 2, 3), 0) # zero-length vectors are possible too
```
### 2.1.4. Generating arithmetic progressions with `seq` and `:`

We can call the `seq` function to create a sequence of equally-spaced numbers on a linear scale, i.e., an arithmetic progression.

```{r}
seq(from=1, to=15, by=2)
seq(1, 5, length.out=7)
```
Arithmetic progressions with steps equal to 1 or -1 can also be generated via the `:` operator.

```{r}
1:10    # seq(1, 10) or seq(1, 10, 1)
-1:10   # seq(-1, 10) or seq(-1, 10, 1)
-1:-10  # seq(-1, -10) or seq(-1, -10, -1)
```
## Exercise 2.1

```{r}
seq(1, 10, 2)
seq_along(along.with=c(2,3,7)) # is equivalent to
seq(1, 3, 1)
1:3
seq_len(length.out=4) # is equivalent to
seq(1, 4, 1)
1:4
```
We can get by without the `seq_along` and `seq_len` functions, but they are very fast primitives for specific cases.

### 2.1.5. Generating pseudorandom numbers

Generate pseudorandom numbers:
```{r}
runif(7) # uniform U(0, 1)
rnorm(7) # normal N(0, 1)
```

Sampling:

```{r}
sample(1:10, 20, replace=TRUE)  # 20 with replacement (allow repetitions)
sample(1:10, 5, replace=FALSE)  # 5 without replacement (do not repeat)
```
The distribution of the sampled values does not need to be uniform:

```{r}
sample(x=0:1, 20, replace=TRUE, prob=c(0.9, 0.1)) # the length of x and prob must be the same
```
By setting the seed of the random number generator, i.e., resetting its state to a given one, we can obtain results that are _reproducible_.

```{r}
set.seed(12345) # seeds are specified with integers
sample(1:10, 5, replace=TRUE)
sample(1:10, 5, replace=TRUE)
sample(1:10, 5, replace=TRUE) 
```

Setting the seed to the one used previously gives:

```{r}
set.seed(12345)
sample(1:10, 5, replace=TRUE)
```
```{r}
set.seed(12345)
sample(1:10, 10, replace=TRUE)
```
### 2.1.6. Reading data with `scan`

We can invoke the scan function to read all the inputs and convert them to a single numeric vector:

```{r}
scan(paste0("https://github.com/gagolews/teaching-data/raw/",
    "master/marek/euraud-20200101-20200630.csv"), comment.char="#")
```

We used the paste0 function to concatenate two long strings (too long to fit a single line of code) and form a single URL.

## Exercise 2.2

```{r}
help("scan")
```

## Example 2.3

```{r}
plot(scan(paste0("https://github.com/gagolews/teaching-data/raw/",
    "master/marek/euraud-20200101-20200630.csv"), comment.char="#"),
    main="Exchange rates", xlab="Day", ylab="EUR/AUD")
```









